---
title: "Ref/Stress Bug Sites Wide Filter"
author: "Jason Hill"
date: "`r Sys.Date()`"
output: html_document

---

```{r setup, include=FALSE, message = FALSE, warning = FALSE}
knitr::opts_chunk$set(echo = FALSE)

```

```{r message = FALSE, warning = FALSE}
library(readxl)
library(magrittr)
library(tidyverse)
library(tidyr)
library(DT)
library(sf)
#library(inlmisc)
library(leaflet)

# Set working directory

# setwd('C:/Users/ktq89598/Desktop/FishBugIBI/Data/RefStressFilters')
```


# Overview

This script was built using R 4.0.2 (Taking Off Again). And rebuilt under R 4.3.0 (Already Tomorrow) to have markdown run more efficiently.

This tool filters reference and stress sites from Virginia and West Virginia EDAS databases, in this script we cast a "wide net" net on all the data (n=5292). This initial filter looks at 6 stress parameters that were available for the most sites in the database(s). 

This is just step one of two step process, VDEQ and WVDEP biologists will review the results and help refine the reference and stress sites based on their site specific knowledge. WV DEP data was collected between 1999-2017, VDEQ data collections occurred from 2008-2020, but the majority was collected between 2012-2019.

Below is a breakdown of the sites that were available:

    + Total Sites Filtered = 5,292
    + VDEQ Sites = 3,279
    + WVDEP Sites = 2,013
    + Coastal Bioregion = 334
    + Piedmont Bioregion = 1,723
    + Mountain = 3,235


Least Disturbed (Ref) Filter Parameters:

1.	DO >8 mg/L
2.	pH >6,<9
3.	Sp Cond < 250 uS/cm
4.	TN <= 1 mg/L
5.	TP <= 0.02 mg/L
6.	Total Habitat > 150

Sites are assigned R1, R2, R3, or R4 based on the number parameters the site meets.

* R1 = Great meets all ref criteria (6) 
* R2 = Good meets most and of the ref criteria (5)
* R3 = OK meets (4) ref criteria  
* R4 = Sites with very few ref conditions (1-3)

Stress Filters Parameters

1.	DO < 7 mg/L
2.	pH <6,>9
3.	Sp Cond > 500 uS/cm
4.	TN >= 2 mg/L
5.	TP >= 0.05 mg/L
6.	Total Habitat <= 120

Sites are assigned S1, S2, or S3 based on the number parameters the site meets.

* S1 = Most disturbed sites, exceeded multiple stress filters (3+) 
* S2 = Moderately stress sites, exceeded one or two stress filters (1-2) 
* S3 = Counts sites with 0 Stress Parameters

********
# Results

The first step of the process is to load all the sites and stress parameters into R. You can look at the first few rows below, note their are 17 columns at this point in the process. 

```{r, warning=FALSE, allRef.1}
# Load data
RefStressData <- read_excel('data/All_VDEQ_Data.xlsx', sheet='FinalDataForFilter')

# No filtering needed on this sheet, but going with "Select" so no recoding
RefStressDataSelect <- RefStressData

#datatable(RefStressDataSelect, rownames = F, escape= F, extensions = 'Buttons',
#                options = list(dom = 'Bift', scrollX= TRUE, scrollY = '300px',
#                               pageLength = nrow(RefStressDataSelect), buttons=list('copy','colvis', 'excel')))

#head(RefStressDataSelect)
```

After loading the biology sites with stressor paramaters, we then assign the sites a "Ref", "Stress", or "Fair" status based on each filter parameter as described in the overview. As an example, if a site has a Total Habitat score of 132 is would receive an assignment of "Fair", it would need to score 150 or higher for "Ref" or 120 or lower to be "Stress". After assigning all 5000+ sites a stress category for each parameter, the script counts the number of "Ref", "Stress", or "Fair" for each site. After counting each assignment, the sites receive the S1,S2,S3 or R1,R2,R3,R4 designations. See the first few sites in the table below, you can see the columns have expanded to 36 columns. 

```{r, warning=FALSE, allRef.2}
# Create Dataframe with Stress, Ref and Fair Labeled
RefStressDataSelect$TotHabStress <- cut(RefStressDataSelect$TotHab, c(0,120,150,200), labels=c('Stress','Fair','Ref'))
RefStressDataSelect$TNStress <- cut(RefStressDataSelect$TN, c(0,1,2,100), labels=c('Ref','Fair','Stress'))
RefStressDataSelect$TPStress <- cut(RefStressDataSelect$TP, c(0,0.02,0.05,100), labels=c('Ref','Fair','Stress'))
RefStressDataSelect$DOStress <- cut(RefStressDataSelect$DO, c(0,7,8,100), labels=c('Stress','Fair','Ref'))
RefStressDataSelect$PHStress <- cut(RefStressDataSelect$pH, c(0,6,9,100), labels=c('Fair','Ref','Fair2'))
RefStressDataSelect$SpCondStress <- cut(RefStressDataSelect$SpCond, c(0,250,500,100000), labels=c('Ref','Fair','Stress'))


RefStressDataSelect$CountStress <- rowSums(RefStressDataSelect == 'Stress', na.rm=TRUE) #Counts number of sites with Stress values

RefStressDataSelect$CountRef <- rowSums(RefStressDataSelect == 'Ref', na.rm=TRUE) #Counts number of sites with Ref values

RefStressDataSelect$CountFair1 <- rowSums(RefStressDataSelect == 'Fair', na.rm=TRUE) #Counts number of sites with Fair values

RefStressDataSelect$CountFair2 <- rowSums(RefStressDataSelect == 'Fair2', na.rm=TRUE) #Counts number of sites with Fair values

RefStressDataSelect$CountFair <-RefStressDataSelect$CountFair1 + RefStressDataSelect$CountFair2 #Add Sites with Fair(s) together

RefStressDataSelect <-within(RefStressDataSelect, rm(CountFair1, CountFair2)) #removes interm fair counts from dataframe to be tidy!
#within(df, rm(x, y))

RefStressDataSelect$CountFilterParms <- RefStressDataSelect$CountRef + RefStressDataSelect$CountStress + RefStressDataSelect$CountFair


###Label Stress and Ref Condition
RefStressDataSelect$StressLevel <- cut(RefStressDataSelect$CountStress, c(-1,-0.5,0.5,2.5,20), labels=c('S4','S3','S2','S1'))
RefStressDataSelect$RefLevel <- cut(RefStressDataSelect$CountRef, c(0,3,4,5,20), labels=c('R4','R3','R2','R1'))

#Select Stress and Ref Level
RefStressDataSelect$S1 <- rowSums(RefStressDataSelect == 'S1', na.rm=TRUE) #Counts number of sites with S1 (highest stressor count 3+) 
RefStressDataSelect$S2 <- rowSums(RefStressDataSelect == 'S2', na.rm=TRUE) #Counts number of sites with S2 (high stressor count 1-2)
RefStressDataSelect$S3 <- rowSums(RefStressDataSelect == 'S3', na.rm=TRUE) #Counts number of sites with S3 (0 stressor count)

RefStressDataSelect$R1 <- rowSums(RefStressDataSelect == 'R1', na.rm=TRUE) #Counts number of sites with R1 (high ref count All 6)
RefStressDataSelect$R2 <- rowSums(RefStressDataSelect == 'R2', na.rm=TRUE) #Counts number of sites with R2 (high ref count 5 of 6
RefStressDataSelect$R3 <- rowSums(RefStressDataSelect == 'R3', na.rm=TRUE) #Counts number of sites with R3 (medium ref count 4)
RefStressDataSelect$R4 <- rowSums(RefStressDataSelect == 'R4', na.rm=TRUE) #Counts number of sites with R4 (low ref count 0-3)

#datatable(RefStressDataSelect, rownames = F, escape= F, extensions = 'Buttons',
#                options = list(dom = 'Bift', scrollX= TRUE, scrollY = '300px',
#                               pageLength = nrow(RefStressDataSelect), buttons=list('copy','colvis', 'excel')))

#head(RefStressDataSelect)
```

The next step of the process is to combine the reference and stress count together to find the sites with the highest stress and least stress. As an example, I combine sites with 3 stress or more (S1 sites) and sites with 1-3 reference filter (R4 sites). By adding sites that are S1 and R4, I am looking for sites that have a lot of disturbance and very few reference like qualities. These sites are then labeled "Highest Stress". By combining sites with no stress S3 and sites the meet all reference filters R1, I find the sites with the least disturbance, they are assigned "BestRef". I repeat combinations of the reference and stress results until all sites receive one of the eight categories. You can see that we now have 61 columns in the table.   

```{r, warning=FALSE, allRef.3}
###Highest Stress Count (S1 + R4) - 3+ Stressors and Low Reference Count

RefStressDataSelect$HighestStressCount <-RefStressDataSelect$S1 + RefStressDataSelect$R4 #Add Sites with Stress Together
RefStressDataSelect$HighestStress <- cut(RefStressDataSelect$HighestStressCount, c(-0.5,1.5,20), labels=c('','Highest Stress'))

###High Stress Count (S1 + R3) - 3+ Stressors and Med Reference Count

RefStressDataSelect$HighStressCount <-RefStressDataSelect$S1 + RefStressDataSelect$R3 #Add Sites with Stress Together
RefStressDataSelect$HighStress <- cut(RefStressDataSelect$HighStressCount, c(-0.5,1.5,20), labels=c('','High Stress'))

###Moderate Stress Count (S2 + R4) - 1 or 2 Stressors and Low Reference Count

RefStressDataSelect$ModStressCount <-RefStressDataSelect$S2 + RefStressDataSelect$R4 #Add Sites with Stress Together
RefStressDataSelect$ModStress <- cut(RefStressDataSelect$ModStressCount, c(-0.5,1.5,20), labels=c('','Moderate Stress'))

###Moderate2 Stress Count (S2 + R3) - 1 or 2 Stressors and Med Reference Count

RefStressDataSelect$Mod2StressCount <-RefStressDataSelect$S2 + RefStressDataSelect$R3 #Add Sites with Stress Together
RefStressDataSelect$Mod2Stress <- cut(RefStressDataSelect$Mod2StressCount, c(-0.5,1.5,20), labels=c('','Moderate2 Stress'))


###Low Stress Count (S3 + R4) - 0 Stressors but Low Reference Count

RefStressDataSelect$LowStressCount <-RefStressDataSelect$S3 + RefStressDataSelect$R4 #Add Sites with Highest Stress Together
RefStressDataSelect$LowStress <- cut(RefStressDataSelect$LowStressCount, c(-0.5,1.5,20), labels=c('','zLow Stress'))

###Lowest Stress//Weak Ref Count (S3 + R3) - 0 Stressors and Middle Reference Count, Weak Best Available Site Also...

RefStressDataSelect$LowestStressCount <-RefStressDataSelect$S3 + RefStressDataSelect$R3 #Add Sites with Highest Stress Together
RefStressDataSelect$LowestStress <- cut(RefStressDataSelect$LowestStressCount, c(-0.5,1.5,20), labels=c('','zLowest Stress'))


###Best Ref Sites Count (R1 + S3) - 9+ Ref Critera Met and Zero Stressor Count

RefStressDataSelect$BestRefCount <-RefStressDataSelect$R1 + RefStressDataSelect$S3 #Add Sites with Ref and Stress Together
RefStressDataSelect$BestRef <- cut(RefStressDataSelect$BestRefCount, c(-0.5,1.5,20), labels=c('','BestRef'))

###Good Ref Sites Count (R2 + S3) - 7-9 Ref Critera Met and Zero Stressor Count

RefStressDataSelect$GoodRefCount <-RefStressDataSelect$R2 + RefStressDataSelect$S3 #Add Sites with Ref and Stress Together
RefStressDataSelect$GoodRef <- cut(RefStressDataSelect$GoodRefCount, c(-0.5,1.5,20), labels=c('','GoodRef'))

###which stressor is causing the problem with having 11+ ref sites but having one stressor (R1 + S2)
###Best Ref Sites Count (R1 + S2) - 9+ Ref Critera Met and moderate Stressor Count
RefStressDataSelect$BestRefModCount <-RefStressDataSelect$R1 + RefStressDataSelect$S2 #Add Sites with Ref and Stress Together
RefStressDataSelect$BestRefMod <- cut(RefStressDataSelect$BestRefModCount, c(-0.5,1.5,20), labels=c('','BestRefModStress'))

###Best Ref Sites Count (R1 + S1) - 9+ Ref Critera Met and high Stressor Count
RefStressDataSelect$BestRefModCount2 <-RefStressDataSelect$R1 + RefStressDataSelect$S1 #Add Sites with Ref and Stress Together
RefStressDataSelect$BestRefMod2 <- cut(RefStressDataSelect$BestRefModCount2, c(-0.5,1.5,20), labels=c('','BestRefModStress2'))


###which stressor is causing the problem with having 9+ ref sites but having one stressor (R2 + S2)
###Good Ref Sites Count (R2 + S2) - 9+ Ref Critera Met and moderate Stressor Count
RefStressDataSelect$GoodRefModCount <-RefStressDataSelect$R2 + RefStressDataSelect$S2 #Add Sites with Ref and Stress Together
RefStressDataSelect$GoodRefMod <- cut(RefStressDataSelect$GoodRefModCount, c(-0.5,1.5,20), labels=c('','GoodRefModStress'))

###Good Ref Sites Count (R2 + S1) - 9+ Ref Critera Met and highest Stressor Count
RefStressDataSelect$GoodRefModCount2 <-RefStressDataSelect$R2 + RefStressDataSelect$S1 #Add Sites with Ref and Stress Together
RefStressDataSelect$GoodRefMod2 <- cut(RefStressDataSelect$GoodRefModCount2, c(-0.5,1.5,20), labels=c('','GoodRefModStress2'))

#put all stress information in one column 
RefStressDataSelect$StressAll <- with(RefStressDataSelect, paste0(GoodRefMod,GoodRefMod2,BestRefMod2,BestRefMod,HighestStress,HighStress,ModStress,Mod2Stress,LowStress,LowestStress,BestRef,GoodRef))

#datatable(RefStressDataSelect, rownames = F, escape= F, extensions = 'Buttons',
#                options = list(dom = 'Bift', scrollX= TRUE, scrollY = '300px',
#                               pageLength = nrow(RefStressDataSelect), buttons=list('copy','colvis', 'excel')))

#head(RefStressDataSelect)
```

Next we test the category assignment results by a subset of sites that have VSCI readily available (no WEST VADEP sites or Coastal Plain sites). A description of eight stress categories are found below the graph.

```{r, warning=FALSE, fig.width=10,fig.height=9, allREf.4}
# Load data with VSCI scores
RefStressDataPlot <- read_excel('data/RefStressData.xlsx', sheet='VSCI_Check_Final')


#qplot(data=RefStressDataSelect,x=StressAll,y=VSCIVCPMI,geom='boxplot') #boxplot to check assumptions

#Make GGPLOT
###make a another fancy pretty boxplot of all data in ggplot
pAll <- ggplot(data=RefStressDataPlot,aes(x=StressAll,y=VSCI)) +
  geom_boxplot(aes(color=StressAll)) +
  ggtitle('VSCI by RefStress Classification') +
  labs(x='RefStress Classification',y='VSCI') +
  theme(plot.title = element_text(color="red", size=14, face="bold.italic"),
        axis.title.x = element_text(color="black", size=14, face="bold"),
        legend.position="bottom") +
  geom_hline(yintercept=60,linetype = "longdash") ##dash line of passing cutoff
pAll
```

Reference and stress classification.

* Best Ref (n=728) = Best Ref Sites Count (R1 + S3) - 6 Ref Criteria Met and Zero Stressor Count 
* Good Ref (n=1153) = Good Ref Sites Count (R2 + S3) - 5 Ref Criteria Met and Zero Stressor Count 
* GoodRefModStress (n=293) = Good Ref Sites Count (R2 + S2) - 5 Ref Criteria Met but has at least one Stressor
* Highest Stress (n=140) = Most disturbed sites, Highest Stress Count (S1 + R4) - 3+ Stressors and Low Reference Count
* Moderate Stress (n=925) = Moderately stress sites, Moderate Stress Count (S2 + R4) - 1 or 2 Stressors and Low Reference Count
* Moderate2 Stress (n=805) = Moderate2 Stress Count (S2 + R3) - 1 or 2 Stressors and Med Reference Count
* zLow Stress (n=337) = Low Stress Count (S3 + R4) - 0 Stressors but Low Reference Count
* zLowest Stress (n=910) = Lowest Stress/Weak Ref Count (S3 + R3) - 0 Stressors and 4 Reference Filters Met

The histogram below showing the final count for each category of stress.

```{r, warning=FALSE, fig.width=10,fig.height=9, subMets.2}
# basic histogram of filter results by stress
p <- ggplot(RefStressDataPlot, aes(x=StressAll)) + 
  geom_histogram(stat="count")

p

```

# Map of Sites by Reference and Stress Classifications

```{r category thresholds}
RefStressDataSelect_sf <- mutate(RefStressDataSelect, DOlevel = as.factor(case_when(DO > 0 ~ 'HIGHDO'))) %>%
  st_as_sf(coords = c("LongitudeDD", "LatitudeDD"),  # make spatial layer using these columns
           remove = F, # dont remove these lat/lon cols from df
           crs = 4326) 

#TNsummary <- filter(stationSummary_sf, !is.na(medianTN)) 
#TPsummary <- filter(stationSummary_sf, !is.na(medianTP))

BestRef <- filter(RefStressDataSelect_sf, StressAll == 'BestRef' )
GoodRef <- filter(RefStressDataSelect_sf, StressAll == 'GoodRef' )
GoodRefModStress <- filter(RefStressDataSelect_sf, StressAll == 'GoodRefModStress' )
HighestStress <- filter(RefStressDataSelect_sf, StressAll == 'Highest Stress' )
ModerateStress <- filter(RefStressDataSelect_sf, StressAll == 'Moderate Stress' )
ModerateStress2 <- filter(RefStressDataSelect_sf, StressAll == 'Moderate2 Stress' )
LowStress <- filter(RefStressDataSelect_sf, StressAll == 'zLow Stress' )
LowestStress <- filter(RefStressDataSelect_sf, StressAll == 'zLowest Stress' )
#DOtest <- filter(RefStressDataSelect_sf, DOlevel == 'HIGHDO' )
#mediumTN <- filter(stationSummary_sf,  TNlevel == 'Medium TN')
#highTN <- filter(stationSummary_sf,  TNlevel == 'High TN')
#lowTP <- filter(stationSummary_sf,  TPlevel == 'Low TP')
#mediumTP <- filter(stationSummary_sf,  TPlevel == 'Medium TP')
#highTP <- filter(stationSummary_sf,  TPlevel == 'High TP')
```

```{r, out.width = '100%'}
#https://color-hex.org/color/3481eb to get color choices

leaflet(data = c("BestRef","GoodRef","GoodRefModStress","HighestStress","ModerateStress",
                 "ModerateStress2","LowStress","LowestStress")) %>% 
  addTiles() %>% setView(-79.1001, 37.7001, zoom=7) %>% 

  addCircleMarkers(data = BestRef, lng = ~LongitudeDD, lat = ~LatitudeDD,
                   color='black', fillColor= '#2E74D3', fillOpacity = 0.5, stroke=0.1, radius = 4,
                   weight = 2, group="Best Reference Sites",
                   label = ~StationID, layerId = ~StationID,
                   popup = leafpop::popupTable(BestRef, zcol=c('StationID'))) %>%
  addCircleMarkers(data = GoodRef, lng = ~LongitudeDD, lat = ~LatitudeDD,
                   color='black', fillColor= '#3aeb34', fillOpacity = 0.5, stroke=0.1, radius = 4,
                   weight = 2, group="Good Reference Sites",
                   label = ~StationID, layerId = ~StationID,
                   popup = leafpop::popupTable(GoodRef, zcol=c('StationID'))) %>%
  addCircleMarkers(data = GoodRefModStress, lng = ~LongitudeDD, lat = ~LatitudeDD,
                  color='black', fillColor= '#7a49a5', fillOpacity = 0.5, stroke=0.1, radius = 4,
                   weight = 2, group="Good Reference Sites with One Stress",
                   label = ~StationID, layerId = ~StationID,
                   popup = leafpop::popupTable(GoodRefModStress, zcol=c('StationID'))) %>%
  addCircleMarkers(data = HighestStress, lng = ~LongitudeDD, lat = ~LatitudeDD,
                   color='black', fillColor= '#f42c40', fillOpacity = 0.5, stroke=0.1, radius = 4,
                   weight = 2, group="Highest Stress Sites",
                   label = ~StationID, layerId = ~StationID,
                   popup = leafpop::popupTable(HighestStress, zcol=c('StationID'))) %>%
  addCircleMarkers(data = ModerateStress, lng = ~LongitudeDD, lat = ~LatitudeDD,
                   color='black', fillColor= '#f4912c', fillOpacity = 0.5, stroke=0.1, radius = 4,
                   weight = 2, group="Moderate High Stress Sites",
                   label = ~StationID, layerId = ~StationID,
                   popup = leafpop::popupTable(ModerateStress, zcol=c('StationID'))) %>%
  addCircleMarkers(data = ModerateStress2, lng = ~LongitudeDD, lat = ~LatitudeDD,
                   color='black', fillColor= '#FAD3AA', fillOpacity = 0.5, stroke=0.1, radius = 4,
                   weight = 2, group="Moderate Stress Sites (Mod2)",
                   label = ~StationID, layerId = ~StationID,
                   popup = leafpop::popupTable(ModerateStress2, zcol=c('StationID'))) %>%
  addCircleMarkers(data = LowStress, lng = ~LongitudeDD, lat = ~LatitudeDD,
                   color='black', fillColor= '#f3f447', fillOpacity = 0.5, stroke=0.1, radius = 4,
                   weight = 2, group="No Stress No Ref All Fair (Low Stress)",
                   label = ~StationID, layerId = ~StationID,
                   popup = leafpop::popupTable(LowStress, zcol=c('StationID'))) %>%
  addCircleMarkers(data = LowestStress, lng = ~LongitudeDD, lat = ~LatitudeDD,
                   color='black', fillColor= '#AAAA31', fillOpacity = 0.5, stroke=0.1, radius = 4,
                   weight = 2, group="No Stress Some Ref Most Fair (Lowest Stress)",
                   label = ~StationID, layerId = ~StationID,
                   popup = leafpop::popupTable(LowestStress, zcol=c('StationID'))) %>%

   addLayersControl(overlayGroups = c("Best Reference Sites","Good Reference Sites",
                                     "Good Reference Sites with One Stress","Highest Stress Sites",
                                     "Moderate High Stress Sites","Moderate Stress Sites (Mod2)",
                                     "No Stress No Ref All Fair (Low Stress)","No Stress Some Ref Most Fair (Lowest Stress)"),
                   options=layersControlOptions(collapsed=T),
                   position='topleft')

###Old Leaflet Code w/inlmisc package
#https://color-hex.org/color/3481eb to get color choices
#CreateWebMap(maps = c("Topo","Imagery","Hydrography"), collapsed = TRUE, 
#                  options= leafletOptions(zoomControl = TRUE,minZoom = 3, maxZoom = 20,
#                                           preferCanvas = TRUE)) %>%
#  setView(-79.1, 37.7, zoom=7)  %>% 
#  addCircleMarkers(data = BestRef,
#                   color='black', fillColor= '#2E74D3', fillOpacity = 0.5, stroke=0.1, radius = 4,
#                   weight = 2, group="Best Reference Sites",
#                   label = ~StationID, layerId = ~StationID,
#                   popup = leafpop::popupTable(BestRef, zcol=c('StationID'))) %>%
#  addCircleMarkers(data = GoodRef,
#                   color='black', fillColor= '#3aeb34', fillOpacity = 0.5, stroke=0.1, radius = 4,
#                   weight = 2, group="Good Reference Sites",
#                   label = ~StationID, layerId = ~StationID,
#                   popup = leafpop::popupTable(GoodRef, zcol=c('StationID'))) %>%
#  addCircleMarkers(data = GoodRefModStress,
#                   color='black', fillColor= '#7a49a5', fillOpacity = 0.5, stroke=0.1, radius = 4,
#                   weight = 2, group="Good Reference Sites with One Stress",
#                   label = ~StationID, layerId = ~StationID,
#                   popup = leafpop::popupTable(GoodRefModStress, zcol=c('StationID'))) %>%
#  addCircleMarkers(data = HighestStress,
#                   color='black', fillColor= '#f42c40', fillOpacity = 0.5, stroke=0.1, radius = 4,
#                   weight = 2, group="Highest Stress Sites",
#                   label = ~StationID, layerId = ~StationID,
#                   popup = leafpop::popupTable(HighestStress, zcol=c('StationID'))) %>%
#  addCircleMarkers(data = ModerateStress,
#                   color='black', fillColor= '#f4912c', fillOpacity = 0.5, stroke=0.1, radius = 4,
#                   weight = 2, group="Moderate High Stress Sites",
#                   label = ~StationID, layerId = ~StationID,
#                   popup = leafpop::popupTable(ModerateStress, zcol=c('StationID'))) %>%
# addCircleMarkers(data = ModerateStress2,
#                   color='black', fillColor= '#FAD3AA', fillOpacity = 0.5, stroke=0.1, radius = 4,
#                   weight = 2, group="Moderate Stress Sites (Mod2)",
#                   label = ~StationID, layerId = ~StationID,
#                   popup = leafpop::popupTable(ModerateStress2, zcol=c('StationID'))) %>%
#  addCircleMarkers(data = LowStress,
#                   color='black', fillColor= '#f3f447', fillOpacity = 0.5, stroke=0.1, radius = 4,
#                   weight = 2, group="No Stress No Ref All Fair (Low Stress)",
#                   label = ~StationID, layerId = ~StationID,
#                   popup = leafpop::popupTable(LowStress, zcol=c('StationID'))) %>%
#  addCircleMarkers(data = LowestStress,
#                   color='black', fillColor= '#AAAA31', fillOpacity = 0.5, stroke=0.1, radius = 4,
#                   weight = 2, group="No Stress Some Ref Most Fair (Lowest Stress)",
#                   label = ~StationID, layerId = ~StationID,
#                   popup = leafpop::popupTable(LowestStress, zcol=c('StationID'))) %>%
#  inlmisc::AddHomeButton(raster::extent(-83.89, -74.80, 36.54, 39.98), position = "topleft") %>%
#  addLayersControl(baseGroups=c("Topo","Imagery","Hydrography"),
#                   overlayGroups = c("Best Reference Sites","Good Reference Sites",
#                                     "Good Reference Sites with One Stress","Highest Stress Sites",
#                                     "Moderate High Stress Sites","Moderate Stress Sites (Mod2)",
#                                     "No Stress No Ref All Fair (Low Stress)","No Stress Some Ref Most Fair (Lowest Stress)"),
#                   options=layersControlOptions(collapsed=T),
#                   position='topleft')
#  
```

# Table of Results

Below is table of reference and stress classification totals by agency, bioregion, and regional office.

```{r, warning=FALSE, summary}
# https://mikoontz.github.io/data-carpentry-week/lesson_joins.html
# Load data
RefStressDataCount <- read_excel('data/RefStressDataFinalWithCoast.xlsx', sheet='RefStressDataWide')

#count all ref/stress data categories
VA_WVA_Count <- RefStressDataCount %>% count(RefStressCode)

#count just wva ref/stress data categories 
WVA <- filter(RefStressDataCount, DataSource == 'WVA_BUG' ) %>% count(RefStressCode) 

#count just va ref/stress data categories 
VA <- filter(RefStressDataCount, DataSource == 'VDEQ_BUG' ) %>% count(RefStressCode) 

#count just bioregion ref/stress data categories 
VA_MTN <- filter(RefStressDataCount, BioRegion == 'Mountain' ) %>% count(RefStressCode)
VA_PIED <- filter(RefStressDataCount, BioRegion == 'Piedmont') %>% count(RefStressCode)
VA_COAST <- filter(RefStressDataCount, BioRegion == 'Coast' ) %>% count(RefStressCode)

#count just region offices ref/stress data categories 
VA_NRO <- filter(RefStressDataCount, Region == 'NRO' ) %>% count(RefStressCode)
VA_PRO <- filter(RefStressDataCount, Region == 'PRO' ) %>% count(RefStressCode)
VA_TRO <- filter(RefStressDataCount, Region == 'TRO' ) %>% count(RefStressCode)
VA_SWRO <- filter(RefStressDataCount, Region == 'SWRO' ) %>% count(RefStressCode)
VA_VRO <- filter(RefStressDataCount, Region == 'VRO' ) %>% count(RefStressCode)
VA_SCRO <- filter(RefStressDataCount, Region == 'SCRO' ) %>% count(RefStressCode)
VA_WCRO <- filter(RefStressDataCount, Region == 'WCRO' ) %>% count(RefStressCode)

#bullshit join code, ask emma how to clean up!
Join <- left_join(VA_WVA_Count, WVA, by = "RefStressCode", copy = FALSE, suffix = c("Total", "WVDEQ"))
Join2 <- left_join(VA, VA_MTN, by = "RefStressCode", copy = FALSE, suffix = c("VDEQ", "Mountain"))          
Join3 <- left_join(VA_PIED, VA_COAST, by = "RefStressCode", copy = FALSE, suffix = c("Piedmont", "Coast"))  
Join4 <- left_join(VA_NRO, VA_PRO, by = "RefStressCode", copy = FALSE, suffix = c("NRO", "PRO"))
Join5 <- right_join(VA_TRO, VA_SWRO, by = "RefStressCode", copy = FALSE, suffix = c("TRO", "SWRO"))
Join6 <- left_join(VA_VRO, VA_SCRO, by = "RefStressCode", copy = FALSE, suffix = c("VRO", "SCRO"))
Join7 <- left_join(VA_WCRO, VA_WVA_Count, by = "RefStressCode", copy = FALSE, suffix = c("WCRO", "Total"))


JoinAll <- left_join(Join, Join2, by = "RefStressCode") %>%
    left_join(., Join3, by = "RefStressCode") %>% 
    left_join(., Join4, by = "RefStressCode") %>% 
    left_join(., Join5, by = "RefStressCode") %>% 
    left_join(., Join6, by = "RefStressCode") %>% 
    left_join(., Join7, by = "RefStressCode") 
  
  
#left_join(x, y, by='Flag') %>%
#                left_join(., z, by='Flag') 

datatable(JoinAll, rownames = F, escape= F, extensions = 'Buttons',
                options = list(dom = 'Bift', scrollX= TRUE, scrollY = '300px',
                               pageLength = nrow(RefStressDataSelect), buttons=list('copy','colvis', 'excel')))


```
# Table of Stress Sites By EcoRegion

Below is table of stress classification totals by ecoregions (VDEQ data only).


```{r, message = FALSE, warning=FALSE, StressSummary}
# Load data
RefStressDataCount <- read_excel('data/RefStressDataFinalWithCoast.xlsx', sheet='RefStressDataWide')

VARefStressDataCount <- filter(RefStressDataCount, DataSource == "VDEQ_BUG")

VAEcoStressCount <- VARefStressDataCount %>%  
  group_by(EcoRegion) %>%
  filter(TotHabStress == "Stress") %>% count(name = 'TotHabStress') %>%
  left_join(VARefStressDataCount %>%  
              group_by(EcoRegion) %>%
              filter(TNStress == "Stress") %>% count(name = 'TNStress') ) %>%
  left_join(VARefStressDataCount %>%  
              group_by(EcoRegion) %>%
              filter(TPStress == "Stress") %>% count(name = 'TPStress') ) %>%
  left_join(VARefStressDataCount %>%  
              group_by(EcoRegion) %>%
              filter(DOStress == "Stress") %>% count(name = 'DOStress') ) %>%
  left_join(VARefStressDataCount %>%  
              group_by(EcoRegion) %>%
              filter(PHStress == "Stress") %>% count(name = 'PHStress') ) %>%
  left_join(VARefStressDataCount %>%  
              group_by(EcoRegion) %>%
              filter(SpCondStress == "Stress") %>% count(name = 'SpCondStress') )


VAEcoStressCount[is.na(VAEcoStressCount)] <- 0


datatable(VAEcoStressCount, rownames = F, escape= F, extensions = 'Buttons',
                options = list(dom = 'Bift', scrollX= TRUE, scrollY = '300px',
                               pageLength = nrow(VAEcoStressCount), buttons=list('copy','colvis', 'excel')))


```



```{r, warning=FALSE, eval = FALSE, allRefLarryVSCO.1}
# Load data
RefStressDataVSCI <- read_excel('data/VSCI_Data_Larry.xlsx', sheet='Data')

# No filtering needed on this sheet, but going with "Select" so no recoding
RefStressDataSelect <- RefStressDataVSCI

# Create Dataframe with Stress, Ref and Fair Labeled
RefStressDataSelect$TotHabStress <- cut(RefStressDataSelect$TotHab, c(0,120,150,200), labels=c('Stress','Fair','Ref'))
RefStressDataSelect$TNStress <- cut(RefStressDataSelect$TN, c(0,1,2,100), labels=c('Ref','Fair','Stress'))
RefStressDataSelect$TPStress <- cut(RefStressDataSelect$TP, c(0,0.02,0.05,100), labels=c('Ref','Fair','Stress'))
RefStressDataSelect$DOStress <- cut(RefStressDataSelect$DO, c(0,7,8,100), labels=c('Stress','Fair','Ref'))
RefStressDataSelect$PHStress <- cut(RefStressDataSelect$pH, c(0,6,9,100), labels=c('Fair','Ref','Fair2'))
RefStressDataSelect$SpCondStress <- cut(RefStressDataSelect$SpCond, c(0,250,500,100000), labels=c('Ref','Fair','Stress'))


RefStressDataSelect$CountStress <- rowSums(RefStressDataSelect == 'Stress', na.rm=TRUE) #Counts number of sites with Stress values

RefStressDataSelect$CountRef <- rowSums(RefStressDataSelect == 'Ref', na.rm=TRUE) #Counts number of sites with Ref values

RefStressDataSelect$CountFair1 <- rowSums(RefStressDataSelect == 'Fair', na.rm=TRUE) #Counts number of sites with Fair values

RefStressDataSelect$CountFair2 <- rowSums(RefStressDataSelect == 'Fair2', na.rm=TRUE) #Counts number of sites with Fair values

RefStressDataSelect$CountFair <-RefStressDataSelect$CountFair1 + RefStressDataSelect$CountFair2 #Add Sites with Fair(s) together

RefStressDataSelect <-within(RefStressDataSelect, rm(CountFair1, CountFair2)) #removes interm fair counts from dataframe to be tidy!
#within(df, rm(x, y))

RefStressDataSelect$CountFilterParms <- RefStressDataSelect$CountRef + RefStressDataSelect$CountStress + RefStressDataSelect$CountFair


###Label Stress and Ref Condition
RefStressDataSelect$StressLevel <- cut(RefStressDataSelect$CountStress, c(-1,-0.5,0.5,2.5,20), labels=c('S4','S3','S2','S1'))
RefStressDataSelect$RefLevel <- cut(RefStressDataSelect$CountRef, c(0,3,4,5,20), labels=c('R4','R3','R2','R1'))

#Select Stress and Ref Level
RefStressDataSelect$S1 <- rowSums(RefStressDataSelect == 'S1', na.rm=TRUE) #Counts number of sites with S1 (highest stressor count 3+) 
RefStressDataSelect$S2 <- rowSums(RefStressDataSelect == 'S2', na.rm=TRUE) #Counts number of sites with S2 (high stressor count 1-2)
RefStressDataSelect$S3 <- rowSums(RefStressDataSelect == 'S3', na.rm=TRUE) #Counts number of sites with S3 (0 stressor count)

RefStressDataSelect$R1 <- rowSums(RefStressDataSelect == 'R1', na.rm=TRUE) #Counts number of sites with R1 (high ref count All 6)
RefStressDataSelect$R2 <- rowSums(RefStressDataSelect == 'R2', na.rm=TRUE) #Counts number of sites with R2 (high ref count 5 of 6
RefStressDataSelect$R3 <- rowSums(RefStressDataSelect == 'R3', na.rm=TRUE) #Counts number of sites with R3 (medium ref count 4)
RefStressDataSelect$R4 <- rowSums(RefStressDataSelect == 'R4', na.rm=TRUE) #Counts number of sites with R4 (low ref count 0-3)

###Highest Stress Count (S1 + R4) - 3+ Stressors and Low Reference Count

RefStressDataSelect$HighestStressCount <-RefStressDataSelect$S1 + RefStressDataSelect$R4 #Add Sites with Stress Together
RefStressDataSelect$HighestStress <- cut(RefStressDataSelect$HighestStressCount, c(-0.5,1.5,20), labels=c('','Highest Stress'))

###High Stress Count (S1 + R3) - 3+ Stressors and Med Reference Count

RefStressDataSelect$HighStressCount <-RefStressDataSelect$S1 + RefStressDataSelect$R3 #Add Sites with Stress Together
RefStressDataSelect$HighStress <- cut(RefStressDataSelect$HighStressCount, c(-0.5,1.5,20), labels=c('','High Stress'))

###Moderate Stress Count (S2 + R4) - 1 or 2 Stressors and Low Reference Count

RefStressDataSelect$ModStressCount <-RefStressDataSelect$S2 + RefStressDataSelect$R4 #Add Sites with Stress Together
RefStressDataSelect$ModStress <- cut(RefStressDataSelect$ModStressCount, c(-0.5,1.5,20), labels=c('','Moderate Stress'))

###Moderate2 Stress Count (S2 + R3) - 1 or 2 Stressors and Med Reference Count

RefStressDataSelect$Mod2StressCount <-RefStressDataSelect$S2 + RefStressDataSelect$R3 #Add Sites with Stress Together
RefStressDataSelect$Mod2Stress <- cut(RefStressDataSelect$Mod2StressCount, c(-0.5,1.5,20), labels=c('','Moderate2 Stress'))


###Low Stress Count (S3 + R4) - 0 Stressors but Low Reference Count

RefStressDataSelect$LowStressCount <-RefStressDataSelect$S3 + RefStressDataSelect$R4 #Add Sites with Highest Stress Together
RefStressDataSelect$LowStress <- cut(RefStressDataSelect$LowStressCount, c(-0.5,1.5,20), labels=c('','zLow Stress'))

###Lowest Stress//Weak Ref Count (S3 + R3) - 0 Stressors and Middle Reference Count, Weak Best Available Site Also...

RefStressDataSelect$LowestStressCount <-RefStressDataSelect$S3 + RefStressDataSelect$R3 #Add Sites with Highest Stress Together
RefStressDataSelect$LowestStress <- cut(RefStressDataSelect$LowestStressCount, c(-0.5,1.5,20), labels=c('','zLowest Stress'))


###Best Ref Sites Count (R1 + S3) - 9+ Ref Critera Met and Zero Stressor Count

RefStressDataSelect$BestRefCount <-RefStressDataSelect$R1 + RefStressDataSelect$S3 #Add Sites with Ref and Stress Together
RefStressDataSelect$BestRef <- cut(RefStressDataSelect$BestRefCount, c(-0.5,1.5,20), labels=c('','BestRef'))

###Good Ref Sites Count (R2 + S3) - 7-9 Ref Critera Met and Zero Stressor Count

RefStressDataSelect$GoodRefCount <-RefStressDataSelect$R2 + RefStressDataSelect$S3 #Add Sites with Ref and Stress Together
RefStressDataSelect$GoodRef <- cut(RefStressDataSelect$GoodRefCount, c(-0.5,1.5,20), labels=c('','GoodRef'))

###which stressor is causing the problem with having 11+ ref sites but having one stressor (R1 + S2)
###Best Ref Sites Count (R1 + S2) - 9+ Ref Critera Met and moderate Stressor Count
RefStressDataSelect$BestRefModCount <-RefStressDataSelect$R1 + RefStressDataSelect$S2 #Add Sites with Ref and Stress Together
RefStressDataSelect$BestRefMod <- cut(RefStressDataSelect$BestRefModCount, c(-0.5,1.5,20), labels=c('','BestRefModStress'))

###Best Ref Sites Count (R1 + S1) - 9+ Ref Critera Met and high Stressor Count
RefStressDataSelect$BestRefModCount2 <-RefStressDataSelect$R1 + RefStressDataSelect$S1 #Add Sites with Ref and Stress Together
RefStressDataSelect$BestRefMod2 <- cut(RefStressDataSelect$BestRefModCount2, c(-0.5,1.5,20), labels=c('','BestRefModStress2'))


###which stressor is causing the problem with having 9+ ref sites but having one stressor (R2 + S2)
###Good Ref Sites Count (R2 + S2) - 9+ Ref Critera Met and moderate Stressor Count
RefStressDataSelect$GoodRefModCount <-RefStressDataSelect$R2 + RefStressDataSelect$S2 #Add Sites with Ref and Stress Together
RefStressDataSelect$GoodRefMod <- cut(RefStressDataSelect$GoodRefModCount, c(-0.5,1.5,20), labels=c('','GoodRefModStress'))

###Good Ref Sites Count (R2 + S1) - 9+ Ref Critera Met and highest Stressor Count
RefStressDataSelect$GoodRefModCount2 <-RefStressDataSelect$R2 + RefStressDataSelect$S1 #Add Sites with Ref and Stress Together
RefStressDataSelect$GoodRefMod2 <- cut(RefStressDataSelect$GoodRefModCount2, c(-0.5,1.5,20), labels=c('','GoodRefModStress2'))

#put all stress information in one column 
RefStressDataSelect$StressAll <- with(RefStressDataSelect, paste0(GoodRefMod,GoodRefMod2,BestRefMod2,BestRefMod,HighestStress,HighStress,ModStress,Mod2Stress,LowStress,LowestStress,BestRef,GoodRef))

datatable(RefStressDataSelect, rownames = F, escape= F, extensions = 'Buttons',
                options = list(dom = 'Bift', scrollX= TRUE, scrollY = '300px',
                               pageLength = nrow(RefStressDataSelect), buttons=list('copy','colvis', 'excel')))

```

